
---
title: "dataProc"
output: html_notebook
---

required packages:
scde 3.0 (from bioconductor) 
CytobankAPI/tsne (for viSNE)
cyt (for visualizing viSNE) ????
mclust 4.4
igraph 0.7.1
FactoMineR 1.28

```{r}
#scde special install
if (!require("BiocManager", quietly = FALSE))
    install.packages("BiocManager")
```
Run this in terminal before attempting to install scde, because cairo is not pre-installed:
sudo apt-get install libcairo2-dev
```{r}
BiocManager::install("scde")
```

remotes packages can be used to install only mclust, igraph and FactoMineR.
```{r}
#install packages using remotes (with specific version)
install.packages("remotes")  #get and install remotes first
library(remotes) #load remotes before everything else
#install_bioc("3.0/scde",upgrade="never") #installing from bioconductor github page
#install_version("scde","3.0") #not using the bioconductor install function
install_version("mclust","4.4")
install_version("igraph","0.7.1")
install_version("FactoMineR","1.28")
```

```{r}
#install packages without using remotes (no specific version)
#install.packages("bioconductor") #??
#install.packages("scde") 
#install.packages("CytobankAPI") #guaranteed and has ViSNE subpackage, but not sure if it's the one asked by paper (for ViSNE)
install.packages("tsne") #valid package, mentioned in paper but may not be right one because no subpackages (for ViSNE)
install.packages("Rtsne")
#install.packages("mclust")
#install.packages("igraph")
#install.packages("FactoMineR")
```

```{r}
#load packages
#library(CytobankAPI)
library(BiocManager)
library(scde)
library(tsne)
library(Rtsne)
library(mclust)
library(igraph)
library(FactoMineR)
#library(tidyverse) #for tsne
library(knitr) #for outputting and CytobankAPI
```

```{r}
getwd()
```


load data
gse67835
GSE67835_RAW.tar from ncbi
multiple
csv file
/home/jaic1/Documents/srp/gseCSV/GSM1657993_nochipID12.C23.csv
file list:
GSM1657871_1772078217.C03.csv
GSM1657872_1772078217.C04.csv
..
GSM1657990_1772078237.C84.csv
GSM1657991_1772078237.C92.csv
GSM1657992_nochipID12.C05.csv
GSM1657993_nochipID12.C23.csv
..
GSM1658365_nochipID4.C95.csv
GSM1658366_nochipID4.C96.csv



```{r}
#data <- read.csv(file.choose()) #if want to select file from pop-up window
data<-read.csv("../../../gseCSV/GSM1657871_1772078217.C03.csv", header = FALSE, stringsAsFactors = FALSE, sep = "\t") #use relative path
print(data)
```

find a way to put all csv into a single DF/matrix
all input files starts with "GSM165"
all input files has same first column, which acts as index.
suggestion: rename 2nd column's header to match source filename
then, append to bigger dataframe horizontally (by column)

```{r}
df<-read.csv("../../../gseCSV/GSM1657871_1772078217.C03.csv", header = FALSE, stringsAsFactors = FALSE, sep = "\t") #use relative path
filenameList<-c("../../../gseCSV/GSM1657872_1772078217.C04.csv","../../../gseCSV/GSM1657873_1772078217.C06.csv","../../../gseCSV/GSM1657874_1772078217.C07.csv")
fnl<-c("GSM1657872_1772078217.C04","GSM1657873_1772078217.C06","GSM1657874_1772078217.C07")
for (f in fnl){
  fn<-paste("../../../gseCSV/", f,".csv", sep ="") #reconstruct full path and filename
  print(fn)
  df2<-read.csv(fn, header = FALSE, stringsAsFactors = FALSE, sep = "\t")
  #rename second column's header
  df2<-subset(df2,select=-V1)#select only second column, drop first column (deactivate this line to perform all-column append to see if all rows align)
  df<-cbind(df,df2) #append new df
}
  
print(df)
```



how to read this??
GSE67835-GPL18573_series_matrix.txt
GSE67835-GPL15520_series_matrix.txt
```{r}
mat <- read.table("../../../matrix/GSE67835-GPL18573_series_matrix.txt")#, header = TRUE, sep = "")
print(mat)
```


#0. excluded cells that had less than 400000 reads, 
#   reducing the initial dataset of 482 cells to 466


#1. scde: Pairwise distances between cells



#2. viSNE <- tsne: Dimensionality reduction of the distances
#is the source package tsne or Rtsne or cytobankAPI???

##Rtsne method:
#remove missing data
#make sure unique row ID exists

##cytobankAPI method:
#cytobankAPI needs username and password for aunthethication to site???
#viSNE is a Premium Cytobank feature

#endpoint.method -> viSNE , dimensionality_reduction

#copied from https://rdrr.io/cran/CytobankAPI/f/vignettes/cytobank-quickstart.Rmd section: "Making a request"

#Requests to Cytobank API endpoints can only be made after authenticating (see above). 
#The authentication object will be passed as the first parameter of any endpoint request.

# Making a request to the 'experiments' endpoint to create a new experiment
#new_experiment <- experiments.new(cyto_session, experiment_name="My New Experiment", purpose="CytobankAPI quickstart vignette")
#
#View(new_experiment)
#
#c1 <- c(1, 0.007, 0, 0.0023, 0, 0.0026)
#
#new_experiment_dataframe <- data.frame(id=c(1),
#                                       version=c(42),
#                                       purpose=c("CytobankAPI quickstart vignette"),
#                                       comments=c(""),
#                                       public=c(FALSE),
#                                       deleted=c(FALSE),
#                                       sources=c(""),
#                                       experimentName=c("My New Experiment"),
#                                       ...=c("..."),
#                                       stringsAsFactors=FALSE)
#
#knitr::kable(new_experiment_dataframe)


#Generating viSNE maps: (see Supplementary Table 2)
#   2a. uniformly subsample between 6,000 and 12,000 cells


#   2b. ViSNE (500 iterations) -> project data to 2D
#       All runs used:
#           1. an identical random seed 
#           2. default t-SNE parameters: 
#               1. perplexity = 30
#               2. momentum = ...
#                   a. 0.5 for initial 250 iterations
#                   b. 0.8 for remaining (250) iterations
#               3. epsilon = 500
#               4. lie factor = ...
#                   a. 4 for initial 100 iterations
#                   b. 1 for remaining (400) iterations 


#   2c. cyt: visualize viSNE maps and figures 
#            (color coding by subtype, by marker expression levels and in plotting expression level densities).
#            non-existent???


#3. mclust: Subsequent clustering 
#           w/ Bayesian information criterion (BIC) for parameterized Gaussian mixture models fitted by EM algorithm initialized by model-based hierarchical clustering


#4. igraph: Minimum spanning trees, community identification and computation of longest paths 


#5. FactoMineR: Principal components analysis (PCA)

